---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
    message = F, 
    warning = F,
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-", 
    dpi = 100
)
devtools::load_all()
```


# timetk 

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/timetk)](https://cran.r-project.org/package=timetk)
![](http://cranlogs.r-pkg.org/badges/timetk?color=brightgreen)
![](http://cranlogs.r-pkg.org/badges/grand-total/timetk?color=brightgreen)
[![R-CMD-check](https://github.com/business-science/timetk/workflows/R-CMD-check/badge.svg)](https://github.com/business-science/timetk/actions)
[![codecov](https://codecov.io/gh/business-science/timetk/branch/master/graph/badge.svg)](https://app.codecov.io/gh/business-science/timetk)

>  Time series analysis in the `tidyverse`


## Installation

_Download the development version with latest features_:

``` {r, eval = FALSE}
remotes::install_github("business-science/timetk")
```

_Or, download CRAN approved version_:

```{r, eval = FALSE}
install.packages("timetk")
```



## Package Functionality

There are _many_ R packages for working with Time Series data. Here's how `timetk` compares to the "tidy" time series R packages for data visualization, wrangling, and feature engineeering (those that leverage data frames or tibbles). 


<div class = "comparison">

| Task                         | [timetk](https://business-science.github.io/timetk/) | [tsibble](https://tsibble.tidyverts.org/index.html) | [feasts](https://feasts.tidyverts.org/index.html) | [tibbletime](https://business-science.github.io/tibbletime/) |
|------------------------------|--------|---------|---------|-------------|
| __Structure__            |        |         |         |             |             
| Data Structure               | tibble (tbl)  | tsibble (tbl_ts)| tsibble (tbl_ts) | tibbletime (tbl_time)  |
| [__Visualization__](https://business-science.github.io/timetk/articles/TK04_Plotting_Time_Series.html)            |        |         |         |             |             
| Interactive Plots (plotly)   |  ✅  |  :x:  |  :x:  |  :x: | 
| Static Plots (ggplot)        |  ✅  |  :x:  |  ✅   |  :x:  |  
| [Time Series](https://business-science.github.io/timetk/articles/TK04_Plotting_Time_Series.html)                  |  ✅  |  :x:  |  ✅   |  :x:  |  
| [Correlation, Seasonality](https://business-science.github.io/timetk/articles/TK05_Plotting_Seasonality_and_Correlation.html)     |  ✅  |  :x:  |  ✅   |  :x:  |  
| [__Data Wrangling__](https://business-science.github.io/timetk/articles/TK07_Time_Series_Data_Wrangling.html)           |        |         |         |             |             
| Time-Based Summarization     |  ✅  |  :x:  |  :x:   |  ✅  |  
| Time-Based Filtering         |  ✅  |  :x:  |  :x:   |  ✅  |  
| Padding Gaps                 |  ✅  |  ✅  |  :x:   |  :x:  |  
| Low to High Frequency        |  ✅  |  :x:  |  :x:   |  :x:  |  
| Imputation                   |  ✅  |  ✅  |  :x:   |  :x:  |  
| Sliding / Rolling            |  ✅  |  ✅  |  :x:   |  ✅  |  
| __Machine Learning__            |        |         |         |             |
| [Time Series Machine Learning](https://business-science.github.io/timetk/articles/TK03_Forecasting_Using_Time_Series_Signature.html)            |  ✅  |  :x:  |  :x:  |  :x:  |  |
[Anomaly Detection](https://business-science.github.io/timetk/articles/TK08_Automatic_Anomaly_Detection.html)            |  ✅  |  :x:  |  :x:  |  :x:  |  
| [Clustering](https://business-science.github.io/timetk/articles/TK09_Clustering.html)            |  ✅  |  :x:  |  :x:  |  :x:  |  
| [__Feature Engineering (recipes)__](https://business-science.github.io/timetk/articles/TK03_Forecasting_Using_Time_Series_Signature.html)  |        |         |         |             |             
| Date Feature Engineering     |  ✅  |  :x:  |  :x:   |  :x:  |  
| Holiday Feature Engineering  |  ✅  |  :x:  |  :x:   |  :x:  |  
| Fourier Series               |  ✅  |  :x:  |  :x:   |  :x:  |  
| Smoothing & Rolling          |  ✅  |  :x:  |  :x:   |  :x:  |  
| Padding                      |  ✅  |  :x:  |  :x:   |  :x:  |  
| Imputation                   |  ✅  |  :x:  |  :x:   |  :x:  |  
| __Cross Validation (rsample)__  |        |         |         |             |             
| [Time Series Cross Validation](https://business-science.github.io/timetk/reference/time_series_cv.html)      |  ✅  |  :x:  |  :x:   |  :x:  |  
| [Time Series CV Plan Visualization](https://business-science.github.io/timetk/reference/plot_time_series_cv_plan.html) |  ✅  |  :x:  |  :x:   |  :x:  |  
| __More Awesomeness__         |        |         |         |             |             
| [Making Time Series (Intelligently)](https://business-science.github.io/timetk/articles/TK02_Time_Series_Date_Sequences.html)  |  ✅  |  ✅  |  :x:   |  ✅  |  
| [Handling Holidays & Weekends](https://business-science.github.io/timetk/articles/TK02_Time_Series_Date_Sequences.html)        |  ✅  |  :x:  |  :x:   |  :x:  |  
| [Class Conversion](https://business-science.github.io/timetk/articles/TK00_Time_Series_Coercion.html)                    |  ✅  |  ✅  |  :x:   |  :x:  |  
| [Automatic Frequency & Trend](https://business-science.github.io/timetk/articles/TK06_Automatic_Frequency_And_Trend_Selection.html)         |  ✅  |  :x:  |  :x:   |  :x:  |  

</div>


## Getting Started

- [Visualizing Time Series](https://business-science.github.io/timetk/articles/TK04_Plotting_Time_Series.html)

- [Wrangling Time Series](https://business-science.github.io/timetk/articles/TK07_Time_Series_Data_Wrangling.html)

- [Full Time Series Machine Learning and Feature Engineering Tutorial](https://business-science.github.io/timetk/articles/TK03_Forecasting_Using_Time_Series_Signature.html)  

- [API Documentation](https://business-science.github.io/timetk/) for articles and a [complete list of function references](https://business-science.github.io/timetk/reference/index.html). 


## Summary

Timetk is an amazing package that is part of the `modeltime` ecosystem for time series analysis and forecasting. The forecasting system is extensive, and it can take a long time to learn: 

- Many algorithms
- Ensembling and Resampling
- Machine Learning
- Deep Learning
- Scalable Modeling: 10,000+ time series

Your probably thinking how am I ever going to learn time series forecasting. Here's the solution that will save you years of struggling. 

## Take the High-Performance Forecasting Course

> Become the forecasting expert for your organization

<a href="https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting/" target="_blank"><img src="https://www.filepicker.io/api/file/bKyqVAi5Qi64sS05QYLk" alt="High-Performance Time Series Forecasting Course" width="100%" style="box-shadow: 0 0 5px 2px rgba(0, 0, 0, .5);"/></a>

[_High-Performance Time Series Course_](https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting/)

### Time Series is Changing

Time series is changing. __Businesses now need 10,000+ time series forecasts every day.__ This is what I call a _High-Performance Time Series Forecasting System (HPTSF)_ - Accurate, Robust, and Scalable Forecasting. 

 __High-Performance Forecasting Systems will save companies by improving accuracy and scalability.__ Imagine what will happen to your career if you can provide your organization a "High-Performance Time Series Forecasting System" (HPTSF System).

### How to Learn High-Performance Time Series Forecasting

I teach how to build a HPTFS System in my [__High-Performance Time Series Forecasting Course__](https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting). You will learn:

- __Time Series Machine Learning__ (cutting-edge) with `Modeltime` - 30+ Models (Prophet, ARIMA, XGBoost, Random Forest, & many more)
- __Deep Learning__ with `GluonTS` (Competition Winners)
- __Time Series Preprocessing__, Noise Reduction, & Anomaly Detection
- __Feature engineering__ using lagged variables & external regressors
- __Hyperparameter Tuning__
- __Time series cross-validation__
- __Ensembling__ Multiple Machine Learning & Univariate Modeling Techniques (Competition Winner)
- __Scalable Forecasting__ - Forecast 1000+ time series in parallel
- and more.

<p class="text-center" style="font-size:24px;">
Become the Time Series Expert for your organization.
</p>
<br>
<p class="text-center" style="font-size:30px;">
<a href="https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting">Take the High-Performance Time Series Forecasting Course</a>
</p>

## Acknowledgements

The `timetk` package wouldn't be possible without other amazing time series packages. 

* [stats](https://rdrr.io/r/stats/stats-package.html) - Basically every `timetk` function that uses a period (frequency) argument owes it to `ts()`. 
    - `plot_acf_diagnostics()`: Leverages `stats::acf()`, `stats::pacf()` & `stats::ccf()`
    - `plot_stl_diagnostics()`: Leverages `stats::stl()`
* [lubridate](https://lubridate.tidyverse.org/): `timetk` makes heavy use of `floor_date()`, `ceiling_date()`, and `duration()` for "time-based phrases". 
    - Add and Subtract Time (`%+time%` & `%-time%`): `"2012-01-01" %+time% "1 month 4 days"` uses `lubridate` to intelligently offset the day
* [xts](https://github.com/joshuaulrich/xts): Used to calculate periodicity and fast lag automation. 
* [forecast (retired)](https://pkg.robjhyndman.com/forecast/): Possibly my favorite R package of all time. It's based on `ts`, and it's predecessor is the `tidyverts` (`fable`, `tsibble`, `feasts`, and `fabletools`). 
    - The `ts_impute_vec()` function for low-level vectorized imputation using STL + Linear Interpolation uses `na.interp()` under the hood. 
    - The `ts_clean_vec()` function for low-level vectorized imputation using STL + Linear Interpolation uses `tsclean()` under the hood.
    - Box Cox transformation `auto_lambda()` uses `BoxCox.Lambda()`.  
* [tibbletime (retired)](https://business-science.github.io/tibbletime/): While `timetk` does not import `tibbletime`, it uses much of the innovative functionality to interpret time-based phrases:
    - `tk_make_timeseries()` - Extends `seq.Date()` and `seq.POSIXt()` using a simple phase like "2012-02" to populate the entire time series from start to finish in February 2012.
    - `filter_by_time()`, `between_time()` - Uses innovative endpoint detection from phrases like "2012"
    - `slidify()` is basically `rollify()` using `slider` (see below).
* [slider](https://davisvaughan.github.io/slider/): A powerful R package that provides a `purrr`-syntax for complex rolling (sliding) calculations. 
    - `slidify()` uses `slider::pslide` under the hood. 
    - `slidify_vec()` uses `slider::slide_vec()` for simple vectorized rolls (slides).
* [padr](https://edwinth.github.io/padr/): Used for padding time series from low frequency to high frequency and filling in gaps. 
    - The `pad_by_time()` function is a wrapper for `padr::pad()`.
    - See the `step_ts_pad()` to apply padding as a preprocessing recipe!
* [TSstudio](https://github.com/RamiKrispin/TSstudio): This is the best interactive time series visualization tool out there. It leverages the `ts` system, which is the same system the `forecast` R package uses. A ton of inspiration for visuals came from using `TSstudio`. 
