---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
    message = F, 
    warning = F,
    collapse = TRUE,
    comment = "#>",
    fig.path = "README-"
)
```

```{r, echo=FALSE}
library(tidyquant)
library(timekit)
```


# timekit

[![Travis-CI Build Status](https://travis-ci.org/business-science/timekit.svg?branch=master)](https://travis-ci.org/business-science/timekit.svg?branch=master)
[![codecov](https://codecov.io/gh/business-science/timekit/branch/master/graph/badge.svg)](https://codecov.io/gh/business-science/timekit)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/timekit)](https://cran.r-project.org/package=timekit)
![](http://cranlogs.r-pkg.org/badges/timekit?color=brightgreen)
![](http://cranlogs.r-pkg.org/badges/grand-total/timekit?color=brightgreen)

> A collection of tools for working with time series in R

## Benefits

The `timekit` package enables a user to more easily work with time series objects in R. The package has tools for inspecting and manipulating the time-based index (e.g. a date or date time column from a `tbl` or the dates stored as rownames from an `xts` object) and converting time-based objects to and from the many time series classes. The following are key benefits:

* __Index extraction__: get the time series index from any time series object.
* __Understand time series__: create a signature decomposition and summary from a time series index.
* __Build future time series__: create a future time series from an index.
* __Coerce between time-based tibbles (`tbl`) and the major time series data types `xts`, `zoo`, `zooreg`, and `ts`__: Simplifies coercion and maximizes time-based data retention during coercion to regularized time series (e.g. `ts`).


## Tools

The package contains the following functions:

1. __Get an index__: `tk_index` returns the time series index of time
series objects, models. The argument `timekit_idx` can be used to return a
special timekit "index" attribute for regularized `ts` objects that returns a
non-regularized date / date-time index if present.

2. __Get critical timeseries information__: `tk_get_timeseries_signature`
and `tk_get_timeseries_summary` takes an index and provides a time series
decomposition and key summary attributes of the index, respectively. The
`tk_augment_timeseries_signature` expedites adding the time series decomposition to the time
series object. 

3. __Make a future timeseries__: `tk_make_future_timeseries` models a
future time series after an existing time series index.  

4. __Coercion functions__: `tk_tbl`, `tk_ts`, `tk_xts`, `tk_zoo`, and
`tk_zooreg` coerce time-based tibbles `tbl` to and from each of the main
time-series data types `xts`, `zoo`, `zooreg`, `ts`, maintaining the time-based
index.

## Getting started

Load libraries and start with some time series data

```{r}
library(timekit)
library(tidyquant)

FB_tbl <- FANG %>%
    filter(symbol == "FB")
FB_tbl
```

## Work with a time series index

Get the timeseries index

```{r}
idx <- tk_index(FB_tbl)
head(idx)
```

Get the time series signature from the index, a tibble of decomposed features that are useful for understanding the time series observations. 

```{r}
tk_get_timeseries_signature(idx)
```

Get the time series summary from the index, a single-row tibble of key summary information from the time series.

```{r}
tk_get_timeseries_summary(idx)
```

## Make a future time series

Use an index to make a future time series. 

```{r}
holidays <- c("2017-01-02", "2017-01-16", "2017-02-20",
              "2017-04-14", "2017-05-29", "2017-07-04",
              "2017-09-04", "2017-11-23", "2017-12-25") %>%
    ymd()

idx_future <- tk_make_future_timeseries(
    idx, 
    n_future         = 251, 
    skip_values      = holidays, 
    inspect_weekdays = TRUE) 

head(idx_future)
```


## Simplified time series coercion

Coercion to `xts`, `zoo`, or `ts` is simplified. The data is ordered correctly automatically using the column containing the date or datetime information. Non-numeric columns are automatically dropped with a warning to the user (the `silent = TRUE` hides the warnings).

```{r}
# xts
FB_xts <- tk_xts(FB_tbl, silent = TRUE)
```

```{r}
# zoo
FB_zoo <- tk_zoo(FB_tbl, silent = TRUE)
```

```{r}
# ts
FB_ts <- tk_ts(FB_tbl, start = 2013, freq = 252, silent = TRUE)
```

## Maintain time-based index during coercion to regularized classes

Get the default index (regularized for `ts` class).

```{r}
tk_index(FB_ts) %>% head()
```

Get the time-based index using the argument `timekit_idx = TRUE`.

```{r}
tk_index(FB_ts, timekit_idx = TRUE) %>% head()
```


This covers the basics of the `timekit` package capabilities. Here's how to get started.

## Installation

_Download development version with latest features_:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("business-science/timekit")
```

_Or, download CRAN approved version_:

```{r, eval = FALSE}
install.packages("timekit")
```



## Further Information

The `timekit` package includes a vignette to help users get up to speed quickly:

* TK00 - Time Series Coercion Using `timekit`
* TK01 - Working with the Time Series Index using `timekit`

