---
title: "Time Series Coercion Using sweep"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Time Series Coercion Using sweep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
    # message = FALSE,
    # warning = FALSE,
    fig.width = 8, 
    fig.height = 4.5,
    fig.align = 'center',
    out.width='95%', 
    dpi = 200
)
library(tidyquant)
library(sweep)
library(forecast)
# devtools::load_all() # Travis CI fails on load_all()
```

> Simplified and extensible time series coercion tools

The time series landscape in R is vast, deep, and complex causing many inconsistencies in data attributes and formats ultimately making it difficult to coerce between the different data structures. The `zoo` and `xts` packages solved a number of the issues in dealing with the various classes (`ts`, `zoo`, `xts`, `irts`, `msts`, and the list goes on...). However, because these packages deal in classes other than data frame, the issues with coercion between `tbl` and other time series object classes are still present.

The `sweep` package provides tools that solve the issues with coercion, maximizing attribute extensibility (the required data attributes are retained during the coercion to each of the primary time series classes). The following tools are available to coerce and retrieve key information:

* __Coercion functions__: `sw_tbl`, `sw_ts`, `sw_xts`, `sw_zoo`, and `sw_zooreg`. These functions coerce time-based tibbles `tbl` to and from each of the main time-series data types `xts`, `zoo`, `zooreg`, `ts`, maintaining the time-based index.

* __Index function__: `sw_index` returns the index. When the argument, `.sweep_idx = TRUE`, A time-based index (non-regularized index) of `forecast` objects, models, and `ts` objects is returned if present. Refer to `sw_ts()` to learn about non-regularized index persistence during the coercion process.

This vignette includes a brief case study on coercion issues and then a detailed explanation of `sweep` function coercion between time-based `tbl` objects and several primary time series classes (`xts`, `zoo`, `zooreg` and `ts`).

# Prerequisites

Before we get started, load the following packages.

```{r, eval = F}
library(tidyquant)
library(sweep)
```

# Data

We'll use the ten-year treasury rate available from the FRED database with the code, "DGS10". We'll retrieve the data set using `tq_get(get = "economic.data")`. The return structure is a tibble (or "tidy" data frame), which is not conducive to many of the popular time series analysis packages including `quantmod`, `TTR`, `forecast` and many others. 

```{r}
ten_year_treasury_rate_tbl <- tq_get("DGS10", 
                                     get  = "economic.data", 
                                     from = today() - years(20), 
                                     to   = today()) %>%
    rename(pct = price) %>%
    mutate(pct = pct / 100)
ten_year_treasury_rate_tbl
```

For purposes of the Case Study, we'll change to a quarterly periodicity using `tq_transmute()` from the `tidyquant` package. Note that `NA` values are automatically removed from the data (message not shown).

```{r, message=F, warning=F}
ten_year_treasury_rate_tbl <- ten_year_treasury_rate_tbl %>%
    tq_transmute(pct, mutate_fun = to.period, period = "quarters")
ten_year_treasury_rate_tbl
```



# Case Study: Coercion issues with ts()

The `ts` object class has roots in the `stats` package and many popular packages use this time series data structure including the popular `forecast` package. With that said, the `ts` data structure is the most difficult to coerce back and forth because by default it does not contain a time-based index. Rather it uses a regularized index computed using the `start` and `frequency` arguments. Coercion to `ts` is done using the `ts()` function from the `stats` library, which results in various problems.  

## Problems 

First, only numeric columns get coerced. If the user forgets to add the `[,"pct"]` to drop the "date" column, `ts()` returns dates in numeric format which is not what the user wants. 

```{r}
# date column gets coerced to numeric
ts(ten_year_treasury_rate_tbl, start = c(1997, 2), freq = 4) %>%
    head()
```

The correct method is to call the specific column desired. However, this presents a new issue. The date index is lost, and a different "regularized" index is built using the `start` and `frequency` attributes. 

```{r}
ten_year_treasury_rate_ts_stats <- ts(ten_year_treasury_rate_tbl[,"pct"], 
                                      start = c(1997, 2), 
                                      freq  = 4)
ten_year_treasury_rate_ts_stats
```

We can see from the structure (using the `str()` function) that the regularized time series is present, but there is no date index retained.

```{r}
# No date index attribute
str(ten_year_treasury_rate_ts_stats)
```

We can get the index using the `index()` function from the `zoo` package. The index retained is a regular sequence of numeric values. These values cannot be coerced back to the original time-base due to dates and date times containing significantly more information (i.e. year-month-day, hour-minute-second, and timezone attributes).

```{r}
# Regularized numeric sequence
index(ten_year_treasury_rate_ts_stats)
```


## Solution

The `sweep` package contains a new function, `sw_ts()`, that enables maintaining the original date index as an attribute. When we repeat the `tbl` to `ts` coercion process using the new function, `sw_ts()`, we can see a few differences. 

First, only numeric columns get coerced, which prevents unintended consequences due to R coercion rules (e.g. dates getting unintentionally converted or characters causing the homogeneous data structure converting all values to character). If a column is dropped, the user gets a warning. 

```{r}
# date automatically dropped and user is warned
ten_year_treasury_rate_ts_sweep <- sw_ts(ten_year_treasury_rate_tbl, 
                                         start = c(1997, 2), 
                                         freq  = 4)
ten_year_treasury_rate_ts_sweep
```

Second, the data returned has a few additional attributes. The most important of which is a numeric attribute, "index", which contains the original date information as a number. The `ts()` function will not preserve this index while `sw_ts()` will preserve the index in numeric form along with the time zone and class. 

```{r}
# More attributes including time index, time class, time zone
str(ten_year_treasury_rate_ts_sweep)
```

## Advantages of coercion with sw_tbl()

Since we used the `sw_ts()` during coercion, we can extract the original index in date format using `sw_index(.sweep_idx = TRUE)` (the default is FALSE which returns the default regularized index).

```{r}
# Can now retrieve the original date index
index <- sw_index(ten_year_treasury_rate_ts_sweep, .sweep_idx = TRUE)
head(index)
class(index)
```

Next, the `sw_tbl()` function has an argument `.sweep_idx` also which can be used to select which index to return. First, we show coercion using the default index. Notice that the index returned is "regularized" meaning its actually a numeric index rather than a time-based index.

```{r}
# Coercion back to tibble using the default index (regularized)
ten_year_treasury_rate_ts_sweep %>%
    sw_tbl(index_rename = "date", .sweep_idx = FALSE)
```

Now, the benefit of using the `sw_ts()` function is we can now get the contents of the original tibble using the `sw_tbl()` argument `.sweep_idx = TRUE`.

```{r}
# Coercion back to tibble now using the sweep index (date / date-time)
ten_year_treasury_rate_tbl_sweep <- ten_year_treasury_rate_ts_sweep %>%
    sw_tbl(index_rename = "date", .sweep_idx = TRUE)
ten_year_treasury_rate_tbl_sweep
```

We can see that in this case (and in most cases) you can get the same data frame you began with.

```{r}
# Comparing the coerced tibble with the original tibble
identical(ten_year_treasury_rate_tbl_sweep, ten_year_treasury_rate_tbl)
```


# Coercion Methods

Using the `ten_year_treasury_rate_tbl`, we'll go through the various coercion methods using `sw_tbl`, `sw_xts`, `sw_zoo`, `sw_zooreg`, and `sw_ts`.

## From tbl

The starting point is the `ten_year_treasury_rate_tbl`. We will coerce this into `xts`, `zoo`, `zooreg` and `ts` classes.

```{r}
# Start:
ten_year_treasury_rate_tbl
```


### to xts

Use `sw_xts()`. By default "date" is used as the date index and the "date" column is dropped from the output. Only numeric columns are coerced to avoid unintentional coercion issues. 

```{r}
# End
ten_year_treasury_rate_xts <- sw_xts(ten_year_treasury_rate_tbl) 
head(ten_year_treasury_rate_xts)
```

Use the `select` argument to specify which columns to drop. Use the `date_var` argument to specify which column to use as the date index. Notice the message and warning are no longer present.

```{r}
# End - Using `select` and `date_var` args
sw_xts(ten_year_treasury_rate_tbl, select = -date, date_var = date) %>%
    head()
```

Also, as an alternative, we can set `silent = TRUE` to bypass the warnings since the default dropping of the "date" column is what is desired. Notice no messages 

```{r}
sw_xts(ten_year_treasury_rate_tbl, silent = TRUE) %>%
    head()
```


### to zoo

Use `sw_zoo()`. Same as when coercing to xts, the non-numeric "date" column is automatically dropped and the index is automatically selected as the date column. 

```{r}
# End
ten_year_treasury_rate_zoo <- sw_zoo(ten_year_treasury_rate_tbl) 
head(ten_year_treasury_rate_zoo)
```

### to zooreg

Use `sw_zooreg()`. Same as when coercing to xts, the non-numeric "date" column is automatically dropped. The regularized index is built from the function arguments `start` and `freq`.  

```{r}
# End
ten_year_treasury_rate_zooreg <- sw_zooreg(ten_year_treasury_rate_tbl, start = 2011, freq = 12) 
head(ten_year_treasury_rate_zooreg)
```

The original time-based index is retained and can be accessed using `sw_index(.sweep_idx = TRUE)`.

```{r}
# Retrieve original time-based index
sw_index(ten_year_treasury_rate_zooreg, .sweep_idx = TRUE) %>%
    str()
```


### to ts

Use `sw_ts()`. The non-numeric "date" column is automatically dropped. The regularized index is built from the function arguments.  

```{r}
# End
ten_year_treasury_rate_ts <- sw_ts(ten_year_treasury_rate_tbl, start = 2011, freq = 12) 
ten_year_treasury_rate_ts
```

The original time-based index is retained and can be accessed using `sw_index(.sweep_idx = TRUE)`.

```{r}
# Retrieve original time-based index
sw_index(ten_year_treasury_rate_ts, .sweep_idx = TRUE) %>%
    str()
```


## To tbl

Going back to tibble is just as easy using `sw_tbl()`.

### From xts

```{r}
# Start
head(ten_year_treasury_rate_xts)

# End
sw_tbl(ten_year_treasury_rate_xts)
```


### From zoo

```{r}
# Start
head(ten_year_treasury_rate_zoo)

# End
sw_tbl(ten_year_treasury_rate_zoo)
```

### From zooreg

```{r}
# Start
head(ten_year_treasury_rate_zooreg)
```

Index is a regularized numeric sequence.

```{r}
# End - with default index
sw_tbl(ten_year_treasury_rate_zooreg)
```

Index is the original date sequence.

```{r}
# End - with sweep index
sw_tbl(ten_year_treasury_rate_zooreg, .sweep_idx = TRUE)
```


### From ts

```{r}
# Start
ten_year_treasury_rate_ts
```

Index is a regularized numeric sequence.

```{r}
# End - with default index
sw_tbl(ten_year_treasury_rate_ts)
```

Index is the original date sequence.

```{r}
# End - with sweep index 
sw_tbl(ten_year_treasury_rate_ts, .sweep_idx = TRUE)
```

