---
title: "Forecasting Using Multiple Models"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{sweep and dplyr - Forecasting in the tidyverse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
    # message = FALSE,
    # warning = FALSE,
    fig.width = 8, 
    fig.height = 4.5,
    fig.align = 'center',
    out.width='95%', 
    dpi = 200
)
library(tidyquant)
library(sweep)
library(forecast)
# devtools::load_all() # Travis CI fails on load_all()
```

> A "tidy" toolkit for forecasting and time series analysis

One of the most powerful things about `sweep` is that it makes forecasting at scale possible within the "tidyverse". There are two common situations:

1. Applying a model to groups of time series
2. Applying multiple models to a time series

In this vignette we'll review how `sweep` can help the second situation: _Applying multiple models to a time series_.

# Prerequisites

Before we get started, load the following packages.

```{r, eval = F}
library(forecast)
library(tidyquant)
library(sweep)
```


