---
title: "Plotting Time Series"
author: "Matt Dancho"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Plotting Time Series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 8, 
  fig.height = 4.5,
  fig.align = 'center',
  out.width='95%', 
  dpi = 100,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(lubridate)
library(timetk)
```

> timetk: A toolkit for time series analysis in the tidyverse

This tutorial focuses on, `plot_time_series()`, a workhorse time-series plotting function that:

- Generates interactive `plotly` plots (great for exploring & shiny apps)
- Consolidates 20+ lines of `ggplot2` & `plotly` code
- Scales well to many time series
- Can be converted from interactive `plotly` to static `ggplot2` plots


# Plotting Single Time Series

Let's start with a popular time series, `taylor_30_min`, which includes energy demand in megawatts at a sampling interval of 30-minutes. This is a single time series. 

```{r}
taylor_30_min
```

The `plot_time_series()` function generates an interactive `plotly` chart by default. 

- Simply provide the date variable (time-based column) and the numeric variable that changes over time as the first 2 arguments 
- `.plotly_slider = TRUE` adds a date slider to the bottom of the chart. 

```{r}
taylor_30_min %>% plot_time_series(date, value, .plotly_slider = TRUE)
```


# Plotting Groups

Next, let's move on to a dataset with time series groups, `m4_daily`, which is a sample of 4 time series from the M4 competition that are sampled at a daily frequency. 

```{r}
m4_daily %>% group_by(id)
```

Visualizing grouped data is as simple as grouping the data set with `group_by()` prior to piping into the `plot_time_series()` function. Key points:

- Groups can be added in 2 ways: by `group_by()` or by using the `...` to add groups. 
- Groups are then converted to facets. 
- `.facet_ncol = 2` returns a 2-column faceted plot
- `.facet_scales = "free"` allows the x and y-axis of each plot to scale independently of the other plots

```{r}
m4_daily %>%
  group_by(id) %>%
  plot_time_series(date, value, 
                   .facet_ncol = 2, .facet_scales = "free")
```

# Adjusting the Smoother

By default, a LOESS smoother is added to capture trend. This smoother is similar to `geom_smooth()`. However, there are advanced controls including being able to change the flexibility of the smoother using time-based phrases: 

- `.smooth_period = "auto"` Selects 75% of the data for the smoother span. 
- `.smooth_period = 365.5 * 2` Selects exactly 731 observations, or 2-years of data with no m
- `.smooth_period = "2 years"` Uses a trend consisting of 2 full years of data. 
- `.smooth_span = 0.75` Uses 75% of the data. Overrides the `.smooth_period`. 

Note that setting `smooth_message = TRUE` generates the number of observations used for the trend. 

```{r, message=TRUE}
m4_daily %>%
  group_by(id) %>%
  plot_time_series(date, value, 
                   .facet_ncol = 2, .facet_scales = "free", 
                    # Note that D10 and D410 have incomplete windows (a shorter trend is selected)
                   .smooth_period  = "2 year", 
                   .smooth_message = TRUE)     # Shows the trend period selected
```



# Visualizing Transformations & Sub-Groups



```{r}
m4_hourly %>% group_by(id)
```

```{r}
m4_hourly %>%
  plot_time_series(date, log(value),             # Apply a Log Transformation
                   id,                           # Faceting (Group) 
                   .color_var = week(date),      # Color applied to Week transformation
                   # Facet formatting
                   .facet_ncol = 2, 
                   .facet_scales = "free")
```

# Static ggplot2 Visualizations

All of the visualizations can be converted from interactive `plotly` (great for exploring and shiny apps) to static `ggplot2` visualizations (great for reports).

```{r}
taylor_30_min %>%
  plot_time_series(date, value, .interactive = FALSE)
```

# Learning More

If you are interested in learning Time Series Analysis & Forecasting, then I have an Advanced Time Series Course coming soon.


