---
title: "Anomaly Detection"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Anomaly Detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    fig.width = 8, 
    fig.height = 4.5,
    fig.align = 'center',
    out.width='95%', 
    dpi = 100
)

# devtools::load_all() # Travis CI fails on load_all()
```


__Anomaly detection__ is an important part of time series analysis:

1. Detecting anomalies can signify special events
2. Cleaning anomalies can improve forecast error

In this short tutorial, we will cover the `plot_anomaly_diagnostics()` and `tk_anomaly_diagnostics()` functions for visualizing and automatically detecting anomalies at scale. 


```{r}
library(dplyr)
library(purrr)
library(timetk)
```

# Data

This tutorial will use the `walmart_sales_weekly` dataset: 

- Weekly
- Sales spikes at various events 

```{r}
walmart_sales_weekly
```

# Anomaly Visualization

Using the `plot_anomaly_diagnostics()` function, we can interactively detect anomalies at scale. 

```{r, fig.height=7}
walmart_sales_weekly %>%
  group_by(Store, Dept) %>%
  plot_anomaly_diagnostics(Date, Weekly_Sales, .facet_ncol = 2)
```


# Automatic Anomaly Detection

To get the data on the anomalies, we use `tk_anomaly_diagnostics()`, the preprocessing function. 

```{r}
walmart_sales_weekly %>%
  group_by(Store, Dept) %>%
  tk_anomaly_diagnostics(Date, Weekly_Sales)
```


# Learning More

<p>
	<iframe width="100%" height="450" src="https://www.youtube.com/embed/elQb4VzRINg" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen="" style="box-shadow: 0 0 5px 2px rgba(0, 0, 0, .5);"><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span>
	</iframe>
</p>

_My Talk on High-Performance Time Series Forecasting_

Time series is changing. __Businesses now need 10,000+ time series forecasts every day.__ This is what I call a _High-Performance Time Series Forecasting System (HPTSF)_ - Accurate, Robust, and Scalable Forecasting. 

 __High-Performance Forecasting Systems will save companies MILLIONS of dollars.__ Imagine what will happen to your career if you can provide your organization a "High-Performance Time Series Forecasting System" (HPTSF System).

I teach how to build a HPTFS System in my __High-Performance Time Series Forecasting Course__. If interested in learning Scalable High-Performance Forecasting Strategies then [take my course](https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting). You will learn:

- Time Series Machine Learning (cutting-edge) with `Modeltime` - 30+ Models (Prophet, ARIMA, XGBoost, Random Forest, & many more)
- NEW - Deep Learning with `GluonTS` (Competition Winners)
- Time Series Preprocessing, Noise Reduction, & Anomaly Detection
- Feature engineering using lagged variables & external regressors
- Hyperparameter Tuning
- Time series cross-validation
- Ensembling Multiple Machine Learning & Univariate Modeling Techniques (Competition Winner)
- Scalable Forecasting - Forecast 1000+ time series in parallel
- and more.

<p class="text-center" style="font-size:30px;">
<a href="https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting">Unlock the High-Performance Time Series Forecasting Course</a>
</p>
