---
title: "Time Series Clustering"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Time Series Clustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    fig.width = 8, 
    fig.height = 4.5,
    fig.align = 'center',
    out.width='95%', 
    dpi = 100
)

# devtools::load_all() # Travis CI fails on load_all()
```


__Clustering__ is an important part of time series analysis that allows us to organize time series into groups by combining "tsfeatures" (summary matricies) with unsupervised techniques such as K-Means Clustering. In this short tutorial, we will cover the `tk_tsfeatures()` functions that computes a time series feature matrix of summarized information on one or more time series. 

# Libraries

To get started, load the following libraries. 

```{r}
library(dplyr)
library(timetk)
```

# Data

This tutorial will use the `walmart_sales_weekly` dataset: 

- Weekly
- Sales spikes at various events 

```{r}
walmart_sales_weekly
```

# TS Features

Using the `tk_tsfeatures()` function, we can quickly get the "tsfeatures" for each of the time series. A few important points:

- The `features` parameter come from the `tsfeatures` R package. Use one of the function names from `tsfeatures` R package e.g.("lumpiness", "stl_features").

- We can supply any function that returns an aggregation (e.g. "mean" will apply the `base::mean()` function).

- You can supply custom functions by creating a function and providing it (e.g. `my_mean()` defined below)

```{r, fig.height=7}
# Custom Function
my_mean <- function(x, na.rm=TRUE) {
  mean(x, na.rm = na.rm)
}

tsfeature_tbl <- walmart_sales_weekly %>%
    group_by(id) %>%
    tk_tsfeatures(
      .date_var = Date,
      .value    = Weekly_Sales,
      .period   = 52,
      .features = c("frequency", "stl_features", "entropy", "acf_features", "my_mean"),
      .scale    = TRUE,
      .prefix   = "ts_"
    ) %>%
    ungroup()

tsfeature_tbl
```


# Clustering with K-Means

We can quickly add cluster assignments with the `kmeans()` function and some tidyverse data wrangling. 

```{r}
set.seed(123)


cluster_tbl <- tibble(
    cluster = tsfeature_tbl %>% 
        select(-id) %>%
        as.matrix() %>%
        kmeans(centers = 3, nstart = 100) %>%
        pluck("cluster")
) %>%
    bind_cols(
        tsfeature_tbl
    )

cluster_tbl
```

# Visualize the Cluster Assignments

Finally, we can visualize the cluster assignments by joining the `cluster_tbl` with the original `walmart_sales_weekly` and then plotting with `plot_time_series()`. 

```{r}
cluster_tbl %>%
    select(cluster, id) %>%
    right_join(walmart_sales_weekly, by = "id") %>%
    group_by(id) %>%
    plot_time_series(
      Date, Weekly_Sales, 
      .color_var   = cluster, 
      .facet_ncol  = 2, 
      .interactive = FALSE
    )
```


# Learning More

<p>
	<iframe width="100%" height="450" src="https://www.youtube.com/embed/elQb4VzRINg" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen="" style="box-shadow: 0 0 5px 2px rgba(0, 0, 0, .5);"><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span><span id="selection-marker-1" class="redactor-selection-marker"></span>
	</iframe>
</p>

_My Talk on High-Performance Time Series Forecasting_

Time series is changing. __Businesses now need 10,000+ time series forecasts every day.__ This is what I call a _High-Performance Time Series Forecasting System (HPTSF)_ - Accurate, Robust, and Scalable Forecasting. 

 __High-Performance Forecasting Systems will save companies MILLIONS of dollars.__ Imagine what will happen to your career if you can provide your organization a "High-Performance Time Series Forecasting System" (HPTSF System).

I teach how to build a HPTFS System in my __High-Performance Time Series Forecasting Course__. If interested in learning Scalable High-Performance Forecasting Strategies then [take my course](https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting). You will learn:

- Time Series Machine Learning (cutting-edge) with `Modeltime` - 30+ Models (Prophet, ARIMA, XGBoost, Random Forest, & many more)
- NEW - Deep Learning with `GluonTS` (Competition Winners)
- Time Series Preprocessing, Noise Reduction, & Anomaly Detection
- Feature engineering using lagged variables & external regressors
- Hyperparameter Tuning
- Time series cross-validation
- Ensembling Multiple Machine Learning & Univariate Modeling Techniques (Competition Winner)
- Scalable Forecasting - Forecast 1000+ time series in parallel
- and more.

<p class="text-center" style="font-size:30px;">
<a href="https://university.business-science.io/p/ds4b-203-r-high-performance-time-series-forecasting">Unlock the High-Performance Time Series Forecasting Course</a>
</p>
